Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{ Funded by DARPA Cyber Assured Systems Engineering (CASE) }

### 2021-04-23 paper/intro.tex
Add some related work on synthesis
Add draft of component specification

@@

-- Scheduling and Trace Relationships: we assume a stream of correct data-types and the existence of some scheduler to schedule components appropriately. In the later section we give an example of a system that uses seL4 with a pacer to schedule over different domains to provide memory isolation. Scheduling atd the relation to the trace semantics is not part of this paper.

\cite{kalhauge2018sound}

>>>>

Here is the basic process:
\begin{enumerate}
  \item CakeML provides a mechanism to take a defined logic in HOL4 and turn it into CakeML in a way that exactly preserves the meaning of the logic.
  \item Verified components are created by defining the meaning of the components in HOL4, and then using the existing theories to turn that meaning into CakeML
  \item Contiguity types provide a formalism for defining a language and recognizing membership in a language via a matcher
  \item Contiguity types include dependent types and the ability to define when a type is malformed
  \item Define a contract language that makes assumptions on input and guarantees properties on output
  \item Give meaning to the contract language in terms of traces (not easy to do because of the state)
  \item Show how the contract can be expressed in terms of a step function 
  \item The step-function is lifted to traces
  \item A contract defines a language over traces
  \item A step function recognizes membership in that language (prefix closed)
  \item Follow the same pattern used in contiguity types
  \item The CakeML comes for free once the HOL4 theories are shown
\end{enumerate}

@@

Scheduling and Trace Relationships: we assume a stream of correct data-types and the existence of some scheduler to schedule components appropriately. In the later section we give an example of a system that uses seL4 with a pacer to schedule over different domains to provide memory isolation. Scheduling atd the relation to the trace semantics is not part of this paper.

Revised outline:

Describe the system level context with the AGREE analysis that proves system level properties of streams. A system is a collection of components and connections. We assume a finite set of defined datatypes for the connections. A well-formed system is acyclic and provides some suitable set of input streams to consume.

Every component is defined by its inputs, outputs, and a contract. The contract language is simplified AGREE: requires, guarantees, eq, and expressions that include \emph{pre}. Requires are pre-conditions assumed on the input. Guarantees are post-conditions the output must meet. A component consumes a stream of input and produces and stream of output. A well-formed component is also acyclic.

For a component to by synthesized, its contract must be strong enough to define the meaning of every output (e.g., the value assumed by the output and under what conditions it assumes that value).

The AGREE analysis reasons about the system with its component composition proving that all contracts are respected. Failures are counter-examples where some input requirement is violated. The analysis provides a system level guarantee.

Component synthesis compiles the contract to imperative code that implements the contract. The claim is that the final compiled code preserves the meaning of the contract: it produces the same output stream for any input stream that conforms to its pre-condition. The synthesis is accomplished with the CakeML framework that provides the additional guarantee that the post-condition holds over an infinite stream (e.g., that it the component does not terminate and always satisfies the post-condition if the pre-condition holds---Hoare logic extended to statement that diverge).

The semantics is synchronous data-flow. Computation takes zero time as does transport. The system must be well-formed.

Motivate with some example: watchdog example on DNS monitoring for DOS attacks? Estimates the growth rate of the input request queue over time and sends a throttle signal if the growth is too fast (looks suspicious)---any bounded response of behavior based watch-dog is fine.

### 2021-04-23 paper/related-work.tex
Add draft of component specification

>>>>

Lustre langugage definition and semantics \cite{10.1145/41625.41641,97300}. Classical clock directed compilation for synchronous dataflow \cite{10.1145/1379023.1375674,10.1145/2345141.2248426}.

A denotational Kahn semantics for synchronous dataflow is in \cite{10.1007/978-3-540-45212-6_10}.

State machine semantics can be added to synchronous dataflow \cite{10.1145/1086228.1086261}. The idea is to translate imperative constructs into equivalent synchronous dataflow constructs. The resulting Lustre can them be compiled to target platform. The goal is to provide a seamless connection between pure datalflow and pure control design.

There is a fully verified compiler that takes Lustre and turns it into a binary executable \cite{10.1145/3140587.3062358}. The first compilation pass turns synchronous dataflow in Lustre to assembly. The compilation is specified and verified in Coq. The key is in combining infinite sequences of dataflow models with incremental manipulation of memories akin to an imperative model. CompCert is used on the backend to create the final rendered executable.

### 2021-04-23 paper/system.tex
Add draft of component specification

A \emph{system} is a collection of \emph{components}, \emph{connections} between components, a \emph{scheduler} to order execution, and a \emph{system environment} for primary inputs. The computation for the component in this work is defined entirely by a step function:
\[
\konst{stepFn} : \mathit{inports} \times \mathit{stateVars} \to \mathit{stateVars}
\]
The scheduler \emph{activates} components in some order. Activating a component is defined as follows: 
\[
\begin{array}{ll}
 \mathit{inportVals} & = \konst{readInputs}(); \\
 (v_1,\ldots,v_k) & = \konst{readState}() ; \\
 ({v_1}',\ldots,{v_k}') & = \konst{stepFn} (\mathit{inportVals},\mathit{stateVars}) ; \\
 \multicolumn{2}{l}{\konst{writeState}({v_1}',\ldots,{v_k}');} \\
 \multicolumn{2}{l}{\konst{writeOutputs}({v_1}',\ldots,{v_k}');} \\
\end{array}
\]
It is assumed that the scheduler follows some sensible partial order of component activation and allows each component sufficient time for its computation.

>>>>

\begin{figure}
  \[
    \begin{array}{rcl}
      \mathit{c}    & = & \konst{input}\ [(f : \tau)\ldots] \\
                    &   & \konst{output}\ [(f : \tau)\ldots] \\
                    &   & \konst{components}\ [(f : c)\ldots]\\ 
                    &   & \konst{eq}\ [(\lval : \tau := exp) \dots] \\
                    &   & \konst{assume}\ [bexp\ldots] \\
                    &   & \konst{guarantee}\ [bexp\ldots] \\ \\

      \mathit{lval} & = & f \mid \mathit{lval} \, [ \mathit{exp} ]
                          \mid \mathit{lval} . f \\ \\

      f             & = & \mathit{varName} \\ \\

      \mathit{exp}  & = & \konst{Loc}\; \mathit{lval}
                          \mid \konst{nLit}\; \konst{nat}
                          \mid \mathit{constname} \\
                    & | & \mathit{exp} + \mathit{exp}
                          \mid \mathit{exp} * \mathit{exp} \\
                    & | & (exp\ \rightarrow\ exp) \\
                    & | & (\konst{pre}\ exp) \\
                    & | & (\konst{ite}\ bexp\ exp\ exp)\\
                    & | & bexp \\ \\

      \mathit{bexp} & = & \konst{bLoc}\; \mathit{lval}
                          \mid  \konst{bLit}\; \konst{bool}
                          \mid  \neg \mathit{bexp}
                          \mid  \mathit{bexp} \land \mathit{bexp} \\
                    & | & \mathit{exp} = \mathit{exp} 
                    \mid  \mathit{exp} < \mathit{exp}
\end{array}
\]
\caption{Syntax for component specifications.}
\label{fig:syntax}
\end{figure}

A \emph{system} in this model of computation is a \emph{component} that is defined by a specification (see \figref{fig:syntax}). A specification defines inputs, outputs, sub-components, local variables ($\konst{eq}$), assumptions for pre-conditions, and guarantees for post-conditions. A type $\tau$ is a contiguity type which is a self-describing dependent type specification (add citation). An $\mathit{lval}$ is a reference to an \emph{L-value} from compiler concepts and is an expression that can appear on the right hand side of an assignment. Alpha renaming is assumed so that every \emph{lval} is unique.

An environment, $\theta: \mathit{lval} \mapsto \konst{string}$ binds L-values to strings. $\Delta : \konst{string} \to \mathbb{N}$ binds constant names to numbers. Functions $\konst{toN}:\konst{string}\to\mathbb{N}$ and $\konst{toB}:\konst{string}\to\konst{bool}$ interpret byte sequences to numbers and booleans, respectively.

The semantics are synchronous data-flow defined over a sequence of environments where $\theta^i$ is the $i^\mathrm{th}$ environment in the stream. Expression evaluation is defined in the context of the environment stream in \figref{fig:eval}.
\begin{figure*}
\[
\begin{array}{l}
\konst{eval}\; i\; e =
\mathtt{case}\; e\
 \left\{
 \begin{array}{lcl}
    \konst{Loc}\; \lval & \Rightarrow & \konst{toN}(\theta^i(\lval)) \\
    \konst{nLit}\; n & \Rightarrow & n  \\
    \mathit{constname} & \Rightarrow & \Delta(\mathit{constname})  \\
    e_1 + e_2 & \Rightarrow & \konst{eval}\; i \; e_1 + \konst{eval}\; i \; e_2  \\
    e_1 * e_2 & \Rightarrow & \konst{eval}\; i \; e_1 * \konst{eval}\; i \; e_2  \\
    e_1 \rightarrow e_2 & \Rightarrow &  \mathbf{if}\; i = 0\; \mathbf{then}\; \konst{eval}\; i \; e_1\; 
                                         \mathbf{else}\; \konst{eval}\; i \; e_2 \\
    (\konst{pre}\; e) & \Rightarrow &  \konst{eval}\; i-1 \; e
  \end{array}
 \right.
 \\ \\
\konst{evalB}\; i \; b =
\mathtt{case}\; b\
 \left\{
 \begin{array}{lcl}
    \konst{bLoc}\; \lval & \Rightarrow & \konst{toB}(\theta^i(\lval)) \\
    \konst{bLit}\; b & \Rightarrow & b \\
    \neg b & \Rightarrow & \neg(\konst{evalB} \; b)  \\
    b_1 \lor b_2 & \Rightarrow & \konst{evalB}\; i \;b_1 \lor \konst{evalB}\; i \;b_2   \\
    b_1 \land b_2 & \Rightarrow & \konst{evalB}\; i \;b_1 \land \konst{evalB}\; i \;b_2   \\
    e_1 = e_2 & \Rightarrow & \konst{eval} \;e_1 = \konst{eval}\; i \;e_2   \\
    e_1 < e_2 & \Rightarrow & \konst{eval} \;e_1 < \konst{eval}\; i \;e_2
  \end{array}
 \right.
\end{array}
\]
\caption{Expression evaluation in the context of a stream on environments.}
\label{fig:eval}
\end{figure*}

The initial environment stream only contains mappings for primary inputs along the entire stream. Stepping the component updates the current environment. In other words, at the $i^\mathrm{th}$ step, $\theta^i$ is updated with the result of each component evaluation. Stepping a component is recursively defined similar to the eval functions but track the path for the environment binding.
